formula1

### 2019-12-16 inst/formula1.Rmd
more hypotheses

@@

This script takes about a minute to run.  It creates hyper2 objects
`F1_2014` through `F1_2017`, which are the likelihood functions for
the F1 results table for each of the years.

>>>>

(in the diagram above, we remove the bottom-ranked two players, Resta
and Button, whose estimated strength is zero).

@@

We can consider Hamilton and ask various questions about his playing
strength.  One natural null would be that Hamilton's strength is equal
to the mean strength, or $\frac{1}{25}$.

* $H_2\colon p_\mathrm{Hamilton}\geq\frac{1}{25}$
* $H_3\colon p_\mathrm{Hamilton} < \frac{1}{25}$

Noting that the evaluate for Hamilton's strength
$\widehat{p_\mathrm{Hamilton}}$ is about 0.22, we may consider the
likelihood for $H_2$ to be equal to the unconstrained maximum
likelhood as in `L2` above.  The likelihood for $H_3$ requires a
constrained optimization:

```{r,constham}
small <- 0.01
L3 <- maxp(L2017, startp=c(small, rep((1-small)/24,23)),fcm=rbind(c(-1,rep(0,23))),fcv=-1/25,give=TRUE)
L3
```

(note that the constrained maximum attains the constraint of
$p_\mathrm{Hamilton}=\frac{1}{25}$).   The likelihood ratio test gives

```{r,liktestivediff}
L2 - L3$value 
```

or a p-value of

```{r,liktestive}
pchisq(2*(L2 - L3$value),df=1,lower.tail=FALSE)
```

## Null of Hamilton being better than other players

We may now test

* $H_4\colon p_\mathrm{Hamilton} > \max_{i\neq\mathrm{Hamilton}} p_i$

This is quite difficult to evaluate formally but we can evaluate it with no loss in accuracy by observing that the second strongest competitor is Bottas and considering instead

* $H_4'\colon p_\mathrm{Hamilton} > p_\mathrm{Bottas}$.

```{r,L4like}
small <- 0.01
L4 <- maxp(L2017, startp=c(small, rep((1-small)/24,23)),fcm=rbind(c(-1,1,rep(0,22))),fcv=0,give=TRUE)
L4
```

Calculating the support and asymptotic p-values:
```{r,l4val}
L2 - L4$value 
pchisq(2*(L2 - L4$value),df=1,lower.tail=FALSE)
```

this script takes about a minute to run.  It creates hyper2 objects
`F1_2014` through `F1_2017`, which are the likelihood functions for
the F1 results table for each of the years.
